---
title: "Reproducible Research: Peer Assessment 1"
author: Michael Crocco
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
File name is activity.csv and is located in activity.zip
Convert date variable class from factor to date
```{r unzip, cache=TRUE}
activity <- read.table(unz("activity.zip", "activity.csv"), header=T, sep=",")
activity$date <- ad.Date(activity$date, "%Y-%m-%d")

```
## What is mean total number of steps taken per day?
Melt and then re-cast the dataframe taking mean steps per day.
Show daily steps as a histagram
```{r}
library(reshape2)
dailyMelt <- melt(activity, id=date)
dailyTotals <- cast(dailyMelt, date ~ variable, sum, na.rm=T)
stepsPerDay <- dailyTotals$steps
meanSteps <- mean(dailyTotals$steps, na.rm=TRUE)
medianSteps <- median(dailyTotals$steps, na.rm=TRUE)

```
The mean number of steps per day is: `r meanSteps`.
The median number of steps per day is: `r medianSteps`.

Plot histagram with vertical lines for mean and median
```{r histagram, fig.height=4}
hist(stepsPerDay, main="Total Steps per Day")
abline(v=meanSteps, col="green", lty=2)
abline(h=medianSteps, col="red", lty=2)
```
## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
